Titeln på avhandlingens språk

### 2023-02-17 intro.tex
draft

\glspl{WCA} applications are a novel category of wearable, edge-native applications aiming to amplify human cognition in both daily activities and professional settings.
These systems aim to seamlessly integrate into the day-to-day of users, leveraging compute-intensive algorithms to analyze user and environment information and provide real-time, context-aware information and feedback.
\gls{WCA} applications originally emerged as assistive use-cases for individuals suffering from cognitive decline due to aging or traumatic brain injuries~\cite{satyanarayanan2009case,Ha2014towards,Satya2019augmenting}, and have since expanded to a greater range of use cases.
In particular, following the success of non-wearable \gls{XR} and cognitive assistance in industrial settings~\cite{Funk2015Cognitive,Wang2022Comprehensive}, there is increasing interest in the research community in the application of \gls{WCA} for step-by-step assistance for complex assembly tasks~\cite{Chen2017Empirical,belletier2021wearable}.

@@

The combination of these opposing characteristics has led to \gls{WCA} applications being identified in the literature as prime candidates for offloading to the edge~\cite{Ha2014towards,Chen2017Empirical,Chen2018application}.
However, many unknowns still remain before consumer-scale adoption of these applications can become a reality.
One key gap in knowledge pertains to the current lack of tools and methodologies for scalable and repeatable study of \gls{WCA} application performance and resource utilization in realistic deployments.
Due to their human-in-the-loop nature, these applications present a challenge to benchmark and characterize, in particular in real-scale deployments where dozens or even hundreds of users might concurrently use the system.
Accordingly, recruiting a  cohort of subjects for realistic benchmarking and study of \gls{WCA} systems can be prohibitively cumbersome and expensive for many research groups and system designers
There exists therefore a real need for scalable tools for \gls{WCA} benchmarking which do not rely on direct testing of the human-in-the-loop.

@@

In that context, the contributions of this paper are:
\begin{enumerate}
    \item\label{item:contrib:model} We introduce the first, to our knowledge, stochastic model for human timings in \gls{WCA} applications.
    Using the data collected for~\cite{olguinmunoz:impact2021} as a base, we build a stochastic model which takes as input past measurements of system responsiveness and produces realistic step execution times.
    We also introduce a novel way to generate dynamic traces of frames for \gls{WCA} applications which can be combined with the timing model for a full end-to-end emulation of a human.
    We name this new model \emph{\edgedroid}; a direct, more realistic evolution of our initial EdgeDroid approach~\cite{olguin2018scaling,olguin2019edgedroid}.
    \item\label{item:contrib:footprint} Using this model, we study the implications of realistic human behavior for the \emph{application lifetime footprint} of \gls{WCA}, understanding this term as the duration of a specific complete execution of the application task.
    In accordance with previous work~\cite{olguinmunoz:impact2021}, we find dependencies between system responsiveness and human step execution times that lead to substantially different application lifetimes when compared to a first-order baseline which does not take into account human behavior.
    \item\label{item:contrib:optimization} Finally, we study the potential for optimization in \gls{WCA} when considering human behavior using our model.
    We develop a generic model for the stochastic optimization of resource consumption versus responsiveness trade-offs in these applications, which we apply to two potential avenues for \gls{WCA} optimization; number of processed samples and energy consumption per step.
    First, we study the potential for reducing the number of samples captured per step.
    This is a valuable endeavor, as reducing the number of samples captured and subsequently processed directly translates in lower bandwidth demand on the wireless network and processor time demand on the cloudlet.
    Next, we explore the potential for direct optimization of energy consumption.
    The economic feasibility of WCA, and hence its likelihood of commercial adoption depends on it not being a resource hog, and this work furthers that effort.
    Leveraging our more involved user model, combined with the introduced optimization approach, we achieve a \textasciitilde\SI{60}{\percent} reduction in the number of samples processed per step.
    We also achieve an improvement of \SI{20}{\percent} in energy consumption, all while maintaining comparable levels of system responsiveness.
\end{enumerate}

@@

This paper is structured as follows.
\cref{sec:relwork} discusses related work in the field of \gls{WCA} modeling.
In \cref{sec:background} we define and discuss key concepts in \gls{WCA}, as well as summarize the key conclusions from our previous work relating to the relationship between system responsiveness and human behavior in these applications~\cite{olguinmunoz:impact2021}.
In \cref{sec:model} we detail our model for the generation of realistic human timing in \gls{WCA}.
We present its design, verify its expected behavior with respect to our previous results, and introduce the dynamic trace generation for full end-to-end emulation of human behavior.
Next, in \cref{sec:implications:footprint}, we discuss the potential implications of such a model on application lifetimes by studying a small series of representative scenarios.
This is followed by a more in-depth investigation in \cref{sec:implications:optimization} on the potential consequences of a human timing model on the general optimization of \gls{WCA} systems.
In the same section, we introduce our generic optimization framework for resource consumption versus responsiveness trade-offs in \gls{WCA}.
Finally, in \cref{sec:conclusion} we summarize and conclude this paper, as well as briefly discuss potential avenues for future work.

>>>>

\glspl{WCA} applications are a novel category of wearable, edge-native applications aiming to amplify human cognition in both daily activities and professional settings.
These systems aim to seamlessly integrate into the day-to-day of users, leveraging compute-intensive algorithms to analyze user and environment information and provide real-time, context-aware information and feedback.
\gls{WCA} applications originally emerged as assistive use-cases for individuals suffering from cognitive decline due to aging or traumatic brain injuries~\cite{satyanarayanan2009case,ha2014towards,satyanarayanan2019augmenting}, and have since expanded to a greater range of use cases.
In particular, following the success of non-wearable \gls{XR} and cognitive assistance in industrial settings~\cite{funk2015cognitive,wang2022comprehensive}, there is increasing interest in the research community in the application of \gls{WCA} for step-by-step assistance for complex assembly tasks~\cite{chen2017empirical,belletier2021wearable}.

@@

The combination of these opposing characteristics has led to \gls{WCA} applications being identified in the literature as prime candidates for offloading to the edge~\cite{ha2014towards,chen2017empirical,chen2018application}.
However, many unknowns still remain before consumer-scale adoption of these applications can become a reality.
One key gap in knowledge pertains to the current lack of tools and methodologies for scalable and repeatable study of \gls{WCA} application performance and resource utilization in realistic deployments.
Due to their human-in-the-loop nature, these applications present a challenge to benchmark and characterize, in particular in real-scale deployments where dozens or even hundreds of users might concurrently use the system.
Accordingly, recruiting a  cohort of subjects for realistic benchmarking and study of \gls{WCA} systems can be prohibitively cumbersome and expensive for many research groups and system designers
There exists therefore a real need for scalable tools for \gls{WCA} benchmarking which do not rely on direct testing of the human-in-the-loop.

@@

In that context, the contributions of this paper are:
\begin{enumerate}
    \item\label{item:contrib:model} We introduce the first, to our knowledge, stochastic model for human timings in \gls{WCA} applications.
    Using the data collected for~\cite{olguinmunoz2021impact} as a base, we build a stochastic model which takes as input past measurements of system responsiveness and produces realistic step execution times.
    We also introduce a novel way to generate dynamic traces of frames for \gls{WCA} applications which can be combined with the timing model for a full end-to-end emulation of a human.
    We name this new model \emph{\edgedroid}; a direct, more realistic evolution of our initial EdgeDroid approach~\cite{olguinmunoz2018demoscaling,olguinmunoz2019edgedroid}.
    \item\label{item:contrib:footprint} Using this model, we study the implications of realistic human behavior for the \emph{application lifetime footprint} of \gls{WCA}, understanding this term as the duration of a specific complete execution of the application task.
    In accordance with previous work~\cite{olguinmunoz2021impact}, we find dependencies between system responsiveness and human step execution times that lead to substantially different application lifetimes when compared to a first-order baseline which does not take into account human behavior.
    \item\label{item:contrib:optimization} Finally, we study the potential for optimization in \gls{WCA} when considering human behavior using our model.
    We develop a generic model for the stochastic optimization of resource consumption versus responsiveness trade-offs in these applications, which we apply to two potential avenues for \gls{WCA} optimization; number of processed samples and energy consumption per step.
    First, we study the potential for reducing the number of samples captured per step.
    This is a valuable endeavor, as reducing the number of samples captured and subsequently processed directly translates in lower bandwidth demand on the wireless network and processor time demand on the cloudlet.
    Next, we explore the potential for direct optimization of energy consumption.
    The economic feasibility of WCA, and hence its likelihood of commercial adoption depends on it not being a resource hog, and this work furthers that effort.
    Leveraging our more involved user model, combined with the introduced optimization approach, we achieve a \textasciitilde\SI{60}{\percent} reduction in the number of samples processed per step.
    We also achieve an improvement of \SI{20}{\percent} in energy consumption, all while maintaining comparable levels of system responsiveness.
\end{enumerate}

@@

This paper is structured as follows.
\cref{sec:relwork} discusses related work in the field of \gls{WCA} modeling.
In \cref{sec:background} we define and discuss key concepts in \gls{WCA}, as well as summarize the key conclusions from our previous work relating to the relationship between system responsiveness and human behavior in these applications~\cite{olguinmunoz2021impact}.
In \cref{sec:model} we detail our model for the generation of realistic human timing in \gls{WCA}.
We present its design, verify its expected behavior with respect to our previous results, and introduce the dynamic trace generation for full end-to-end emulation of human behavior.
Next, in \cref{sec:implications:footprint}, we discuss the potential implications of such a model on application lifetimes by studying a small series of representative scenarios.
This is followed by a more in-depth investigation in \cref{sec:implications:optimization} on the potential consequences of a human timing model on the general optimization of \gls{WCA} systems.
In the same section, we introduce our generic optimization framework for resource consumption versus responsiveness trade-offs in \gls{WCA}.
Finally, in \cref{sec:conclusion} we summarize and conclude this paper, as well as briefly discuss potential avenues for future work.

### 2023-02-17 model.tex
draft

In the following, we employ the above insights together with the data collected for~\cite{olguinmunoz:impact2021} to design and build a probabilistic model of human behavior for \gls{WCA}.
We detail the construction of a model which uses the data collected to generate, at runtime, realistic execution times. 
In order to accurately emulate the behavior of a human, such a model needs to implement two main behaviors.
First, it needs to generate realistic execution times for each step in the task, considering the current and historical impairment of the \gls{WCA} system, as well as salient individual-difference measures.
We detail this in \cref{ssec:model:exectimes}.
Neuroticism is incorporated in this aspect of the model, as it is the most salient individual difference in the data; however, other factors such as immersive tendency could be treated similarly.
Second, the model must produce sequences of input samples for each step mimicking what a real human would generate; this is explained in \cref{ssec:model:frames}. 
% Finally, in \cref{ssec:model:obtaining} we briefly discuss where implementations of this model can be obtained.

@@

% As expressed in \cref{ssec:plos}, for \textcite{olguinmunoz:impact2021} we collected timing and task performance data for \num{40} participants performing a \num{169}-step task, guided by a \gls{WCA}, as well as personality trait scores for each participant.
% For this paper, we use this data to construct a probabilistic model for the generation of realistic execution times.
The processing of the data from~\cite{olguinmunoz:impact2021} for the generation of execution times can be summarized as a grouping according to discretized levels of neuroticism and a weighted rolling average of \gls{TTF}.
The resulting collections of execution times represent the distributions of these values for users with specific levels of neuroticism, interacting with systems at specific states of impairment and recent histories of impairment.
These distributions can then be sampled to produce new, realistic execution times.
In the following, we detail the step-by-step processing of this data and construction of the probabilistic model.

@@

We verify the behavior of the above described model with respect to the four main conclusions of our work in~\cite{olguinmunoz:impact2021}, which were previously discussed in \cref{ssec:plos}.
These are reformulated as objectives below:

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{publications/2023EdgeDroid2/figs/new_model/ttf_to_exectime.png}
    \caption{%
        Effects of feeding three different \glspl{TTF} (\emph{low}, \SI{0}{\second}, \emph{medium}, \SI{2.5}{\second}, or \emph{high}, \SI{5}{\second}) into the model on the generated execution times.
        Higher \glspl{TTF} directly lead to higher execution times.
        Error bars indicate the \SI{95}{\percent} \gls{CI}.
    }\label{fig:ttf_to_exectime}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{publications/2023EdgeDroid2/figs/new_model/exectime_over_steps.png}
    \caption{%
    Effects of prolonged exposure to constant levels of system impairment on the model.
    At low (\SI{0}{\second}) \gls{TTF}, the models speed-up over time; conversely at high (\SI{5.0}{\second}) \gls{TTF}, the models either present no change or drastically increase their generated execution times, depending on the level of neuroticism.
    Error bands indicate \SI{95}{\percent} \gls{CI}.
    }\label{fig:exectimeduration}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2023EdgeDroid2/figs/new_model/transitions.png}
    \caption{%
        Effects of changes in system impairment on subsequently generated execution times.
        These effects linger on after the change, and thus execution times immediately after a transition are either consistently lower or higher than otherwise at the new \gls{TTF}, depending on the old \gls{TTF}.
        Error bars indicate \SI{95}{\percent} \gls{CI}.
    }\label{fig:transitions}
\end{figure*}

Finally, in \cref{fig:transitions} we showcase the behavior of the model when comparing execution times generated immediately after a change in system responsiveness.
We generate these results by first warming up the model by feeding it a fixed \gls{TTF} (which we will refer to as the \emph{origin} \gls{TTF}) \num{25} times.
Next, another \gls{TTF} value (the \emph{destination} \gls{TTF}) is fed to the model, and we record the output execution time.
Each sample is tagged according to the relation between origin and destination \gls{TTF}, either lower to higher, higher to lower, or equal.
As before, we run \num{600} repetitions of this procedure for each combination of model configuration, origin \gls{TTF}, and destination \gls{TTF}.
Execution times generated immediately after a transition from a higher \gls{TTF} into a lower one are consistently higher than execution times generated without a preceding change in \glspl{TTF}.
Conversely, execution times are consistently lower than otherwise immediately after a change from a lower \gls{TTF} into a higher one.
These results are once again in line with our findings in~\cite{olguinmunoz:impact2021}, in which we found lingering effects of transitions between levels of system impairment on human execution times.

@@

Apart from the aforementioned timing and performance data, for~\cite{olguinmunoz:impact2021} we also recorded all collected video frame samples together with matching metadata.
For each video frame submitted to the \gls{WCA} during the tasks, we recorded
\begin{inlineenum}
    \item the raw video frame captured
    \item sample submission timestamp
    \item \gls{WCA} processing completed timestamp
    \item result or acknowledgement returned timestamp
    \item a tag representing the result of the \gls{WCA} processing
\end{inlineenum}.
The tags assigned corresponded to:
\begin{description}[font={\bfseries\ttfamily}, wide]
    \item[SUCCESS:] frames which triggered a transition to a new step (or the correction of a previous mistake) in the logical task model of the \gls{WCA}, and thus cause the generation of feedback to the user.
    \item[REPEAT:] frames which captured the same board state as the previous successful frame, and thus produced no feedback.
    \item[LOW\_CONFIDENCE] frames for which the image recognition algorithm in the \gls{WCA} did not reach the necessary confidence threshold to interpret it as a valid board state.
        These frames also produce no feedback.
    \item[BLANK] frames in which not enough of the board is visible due to noise, movement, occlusion, etc.
        These frames produce no feedback either.
    \item[TASK\_ERROR] frames which contained an incorrect board state and thus triggered a transition to a procedural corrective step and the generation of feedback to the user.
        However, it must be noted that none of the \num{40} participants made any mistakes during the task, and thus no such frames were encountered in the data.
\end{description}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{publications/2023EdgeDroid2/model_data/frame_probabilities.png}
    \caption{%
        Probability density of frame result tags as a step progresses.
        Note that \texttt{SUCCESS} frames are not included as --- by definition --- the probability for success frames is \num{1} for all normalized instant values greater than or equal to \num{1.0}.
    }\label{fig:frameprobs}
\end{figure}

Using the above insights, together with the corresponding recorded video frames, we devise a scheme for the procedural generation of a synthetic trace for any step in \gls{WCA} task in the same category as those used in~\cite{olguinmunoz:impact2021}.
We first prepare a discretized representation of the probability density map in \cref{fig:frameprobs}.
We segment the normalized instant value into a number of discrete bins (\num{25} in this work), and calculate the relative fraction of frames for each category in each bin.
For each step, given
\begin{inlineenum}
    \item a collection of random non-\texttt{SUCCESS}, non-\texttt{REPEAT} frames (at least one frame for each of the \texttt{BLANK} and \texttt{LOW\_CONFIDENCE} categories)
    \item an appropriate \texttt{SUCCESS} video frame containing the correct state for the step
    \item an appropriate \texttt{REPEAT} video frame containing the correct state for the \emph{previous} step
\end{inlineenum},
we can then procedurally generate a trace by randomly selecting appropriate frames according to the distributions presented in \cref{fig:frameprobs}.
In other words, for each sampling instant in a step with a given execution time \( t_\text{exec} \):
\begin{enumerate}
    \item We calculate \( t_\text{norm} \) according to \cref{eq:tnorm}.
    \item If \( t_\text{norm} \ge 1.0 \), we select the \texttt{SUCCESS} frame and stop sampling.
    \item If instead \( t_\text{norm} < 1.0 \), we find the appropriate bin for \( t_\text{norm} \) and then select a frame by performing a weighted random sampling of the frame categories in the normalized instant bin.
\end{enumerate}

>>>>

In the following, we employ the above insights together with the data collected for~\cite{olguinmunoz2021impact} to design and build a probabilistic model of human behavior for \gls{WCA}.
We detail the construction of a model which uses the data collected to generate, at runtime, realistic execution times. 
In order to accurately emulate the behavior of a human, such a model needs to implement two main behaviors.
First, it needs to generate realistic execution times for each step in the task, considering the current and historical impairment of the \gls{WCA} system, as well as salient individual-difference measures.
We detail this in \cref{ssec:model:exectimes}.
Neuroticism is incorporated in this aspect of the model, as it is the most salient individual difference in the data;
however, other factors such as immersive tendency could be treated similarly.
Second, the model must produce sequences of input samples for each step mimicking what a real human would generate; this is explained in \cref{ssec:model:frames}. 
% Finally, in \cref{ssec:model:obtaining} we briefly discuss where implementations of this model can be obtained.

@@

% As expressed in \cref{ssec:plos}, for \textcite{olguinmunoz:impact2021} we collected timing and task performance data for \num{40} participants performing a \num{169}-step task, guided by a \gls{WCA}, as well as personality trait scores for each participant.
% For this paper, we use this data to construct a probabilistic model for the generation of realistic execution times.
The processing of the data from~\cite{olguinmunoz2021impact} for the generation of execution times can be summarized as a grouping according to discretized levels of neuroticism and a weighted rolling average of \gls{TTF}.
The resulting collections of execution times represent the distributions of these values for users with specific levels of neuroticism, interacting with systems at specific states of impairment and recent histories of impairment.
These distributions can then be sampled to produce new, realistic execution times.
In the following, we detail the step-by-step processing of this data and construction of the probabilistic model.

@@

We verify the behavior of the above described model with respect to the four main conclusions of our work in~\cite{olguinmunoz2021impact}, which were previously discussed in \cref{ssec:plos}.
These are reformulated as objectives below:

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{publications/2023EdgeDroid2/figs/new_model/ttf_to_exectime}
    \caption{%
        Effects of feeding three different \glspl{TTF} (\emph{low}, \SI{0}{\second}, \emph{medium}, \SI{2.5}{\second}, or \emph{high}, \SI{5}{\second}) into the model on the generated execution times.
        Higher \glspl{TTF} directly lead to higher execution times.
        Error bars indicate the \SI{95}{\percent} \gls{CI}.
    }\label{fig:ttf_to_exectime}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[height=25em]{publications/2023EdgeDroid2/figs/new_model/exectime_over_steps}
    \caption{%
    Effects of prolonged exposure to constant levels of system impairment on the model.
    At low (\SI{0}{\second}) \gls{TTF}, the models speed-up over time; conversely at high (\SI{5.0}{\second}) \gls{TTF}, the models either present no change or drastically increase their generated execution times, depending on the level of neuroticism.
    Error bands indicate \SI{95}{\percent} \gls{CI}.
    }\label{fig:exectimeduration}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2023EdgeDroid2/figs/new_model/transitions}
    \caption{%
        Effects of changes in system impairment on subsequently generated execution times.
        These effects linger on after the change, and thus execution times immediately after a transition are either consistently lower or higher than otherwise at the new \gls{TTF}, depending on the old \gls{TTF}.
        Error bars indicate \SI{95}{\percent} \gls{CI}.
    }\label{fig:transitions}
\end{figure*}

Finally, in \cref{fig:transitions} we showcase the behavior of the model when comparing execution times generated immediately after a change in system responsiveness.
We generate these results by first warming up the model by feeding it a fixed \gls{TTF} (which we will refer to as the \emph{origin} \gls{TTF}) \num{25} times.
Next, another \gls{TTF} value (the \emph{destination} \gls{TTF}) is fed to the model, and we record the output execution time.
Each sample is tagged according to the relation between origin and destination \gls{TTF}, either lower to higher, higher to lower, or equal.
As before, we run \num{600} repetitions of this procedure for each combination of model configuration, origin \gls{TTF}, and destination \gls{TTF}.
Execution times generated immediately after a transition from a higher \gls{TTF} into a lower one are consistently higher than execution times generated without a preceding change in \glspl{TTF}.
Conversely, execution times are consistently lower than otherwise immediately after a change from a lower \gls{TTF} into a higher one.
These results are once again in line with our findings in~\cite{olguinmunoz2021impact}, in which we found lingering effects of transitions between levels of system impairment on human execution times.

@@

Apart from the aforementioned timing and performance data, for~\cite{olguinmunoz2021impact} we also recorded all collected video frame samples together with matching metadata.
For each video frame submitted to the \gls{WCA} during the tasks, we recorded
\begin{inlineenum}
    \item the raw video frame captured
    \item sample submission timestamp
    \item \gls{WCA} processing completed timestamp
    \item result or acknowledgement returned timestamp
    \item a tag representing the result of the \gls{WCA} processing
\end{inlineenum}.
The tags assigned corresponded to:
\begin{description}[font={\bfseries\ttfamily}, wide]
    \item[SUCCESS:] frames which triggered a transition to a new step (or the correction of a previous mistake) in the logical task model of the \gls{WCA}, and thus cause the generation of feedback to the user.
    \item[REPEAT:] frames which captured the same board state as the previous successful frame, and thus produced no feedback.
    \item[LOW\_CONFIDENCE] frames for which the image recognition algorithm in the \gls{WCA} did not reach the necessary confidence threshold to interpret it as a valid board state.
        These frames also produce no feedback.
    \item[BLANK] frames in which not enough of the board is visible due to noise, movement, occlusion, etc.
        These frames produce no feedback either.
    \item[TASK\_ERROR] frames which contained an incorrect board state and thus triggered a transition to a procedural corrective step and the generation of feedback to the user.
        However, it must be noted that none of the \num{40} participants made any mistakes during the task, and thus no such frames were encountered in the data.
\end{description}

@@

\begin{figure}
    \centering
    \includegraphics[width=.9\textwidth]{publications/2023EdgeDroid2/model_data/frame_probabilities}
    \caption{%
        Probability density of frame result tags as a step progresses.
        Note that \texttt{SUCCESS} frames are not included as --- by definition --- the probability for success frames is \num{1} for all normalized instant values greater than or equal to \num{1.0}.
    }\label{fig:frameprobs}
\end{figure}

Using the above insights, together with the corresponding recorded video frames, we devise a scheme for the procedural generation of a synthetic trace for any step in \gls{WCA} task in the same category as those used in~\cite{olguinmunoz2021impact}.
We first prepare a discretized representation of the probability density map in \cref{fig:frameprobs}.
We segment the normalized instant value into a number of discrete bins (\num{25} in this work), and calculate the relative fraction of frames for each category in each bin.
For each step, given
\begin{inlineenum}
    \item a collection of random non-\texttt{SUCCESS}, non-\texttt{REPEAT} frames (at least one frame for each of the \texttt{BLANK} and \texttt{LOW\_CONFIDENCE} categories)
    \item an appropriate \texttt{SUCCESS} video frame containing the correct state for the step
    \item an appropriate \texttt{REPEAT} video frame containing the correct state for the \emph{previous} step
\end{inlineenum},
we can then procedurally generate a trace by randomly selecting appropriate frames according to the distributions presented in \cref{fig:frameprobs}.
In other words, for each sampling instant in a step with a given execution time \( t_\text{exec} \):
\begin{enumerate}
    \item We calculate \( t_\text{norm} \) according to \cref{eq:tnorm}.
    \item If \( t_\text{norm} \ge 1.0 \), we select the \texttt{SUCCESS} frame and stop sampling.
    \item If instead \( t_\text{norm} < 1.0 \), we find the appropriate bin for \( t_\text{norm} \) and then select a frame by performing a weighted random sampling of the frame categories in the normalized instant bin.
\end{enumerate}
